Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2021-12-16 src/ms.tex
revising ages discussion
co-author comments

We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages). However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties. In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun, and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.

@@

Recently, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model. Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr. Additionally, a case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}. By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr. The \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however this is sensitive to different \teff scales and different prescriptions for \taucz. We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.

>>>>

We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages). However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties. In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun, and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties. The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and the concentration of stars at the upper age boundary at a given \teff (Figure~\ref{fig:ages}).

@@

Recently, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model. Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr. Additionally, a case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}. By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr. Furthermore, we find that the long-period pile-up lies clearly below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$, which invalidates the claim of \citet{LorenzoOliveira2019}. We note that the \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however, these results are sensitive to different \teff scales and different prescriptions for \taucz. We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.
